Master Thesis

### 2013-06-26 ig.tex
Made small changes in IG (precedence vs types)

We will illustrate the objects and mechanisms of interaction grammars on
a simple example sentence. Since the largest and most-developed
interaction grammar is the Frigram grammar of French, we will be using
French example sentences and their analysis according to
Frigram. Frigram will be discussed in more detail in
subsection~\ref{ssec:frigram}.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.25]{images/ig-partial.pdf}
  \caption{\label{fig:ig-partial} The result of merging the EPTDs of
    \emph{la} and \emph{voit} by merging the nodes (1,5) and (1,5) with
    (2,2) and (2,4), respectively.}
\end{figure}

@@

This example demonstrates two points of interest to us.

@@

Finally, we explain the term $D((F_1 \limp ... \limp F_n \limp
G)^+)$. The $+$ operation recursively assigns positive and negative
polarities to all formula occurrences within $F_1 \limp ... \limp F_n
\limp G$. These polarities then drive the recursive definition of $D$,
which maps this polarized formula into a syntactic description. Since we
will not cover the theorem deeply in our treatise, we will settle for
knowing that the composition $D \circ +$ of the positive polarization
operation $+$ and the syntactic description producing operation $D$ maps
an IILL formula into a syntactic description such that above-stated
theorem holds. Furthermore, this mapping can be easily constructed since
Perrier gives straightforward ways of computing the results of both
operations.

Perrier's theorem establishes a striking symmetry between the two
formalisms. Let us consider the case of parsing a sentence using a
lexicalized grammar in both formalism. First, the sentence is tokenized
into wordforms and for each wordform, a lexical item is selected, be it
a typed constant in some abstract signature for ACGs or an EPTD for IGs.

In ACGs, we will try to take the typed constants and to use them to
produce a term having some distinguished type $S$. Thanks to the
Curry-Howard correspondence, constructing a term having a given type is
nothing more than proving that term's type as a formula in the logic of
our type system. What this means is that our parsing problem boils down
to proving the IILL sequent $\tau_1, ..., \tau_n \vdash S$ (we omit the
inclusion of intuitionistic non-linear implications since they are not
relevant for parsing sentences). The sentence thus becomes parsable by
our grammar under the given lexical selection if and only if this
sequent is provable and furthermore, the proof of the sequent is our
desired syntactic structure in a disguise.

@@

A bigger problem is posed by the fact that $D \circ +$ is not
surjective. Perrier \cite{perrier2001intuitionistic} characterized the
set of syntactic descriptions that can be obtained as images of IILL
sequents using our transformation $D \circ +$, the set of \emph{IILL
  tree descriptions}, which is a proper subset of the set of all
syntactic descriptions. Specifically, in an IILL tree description, all
immediate dominance constraints go from negative nodes to positive
nodes, all (large) dominance constraints lead from positive nodes to
negative nodes and the root has a positive polarity.

@@

The PTDs of IGs include precedence constraints. We can see them as
serving two purposes, to both provide an ordering of the constituents
and to further constrain syntactic composition. In our translation to an
ACG, we could imagine handling the first role, that of providing word
order, by defining a lexicon from our abstract signature. Each constant
would combine its arguments in the order in which their corresponding
subtrees appeared in the ordered PTD. However, the second role of
precedence constraints, to establish further restrictions on syntactic
composition, would be lost, as the only restricting factor on syntactic
composition in ACGs is the type system.

@@

However, this translation is very lossy in terms of its discriminatory
power. Many of the constraints which were forced by the original EPTD
are now ignored. We can imagine improving our strategy by using the
feature structures as our types instead of atomic categories, which
would not only ensure that e.g. the $sent\_type$ of the embedded clause
must be $decl$, but would also capture the information in the polarized
$funct$ feature so that the relative pronoun \emph{que} is not permitted
to extract subjects. There are still constraints that we would miss out
on though, such as the fact that the subject (2,3) of the clause (2,0)
must precede any embedded clauses (2,6)...(2,5), the fact that any such
embedded clause must serve either the $modal$ or the $obj$ function and
the fact that the embedded clause (2,5) must have a subject (2,4).

>>>>

We will illustrate the objects and mechanisms of interaction grammars on
a simple example sentence. Since the largest and most-developed
interaction grammar is the Frigram grammar of French, we will be using
French example sentences and their analyses according to
Frigram. Frigram will be discussed in more detail in
subsection~\ref{ssec:frigram}.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.25]{images/ig-partial.pdf}
  \caption{\label{fig:ig-partial} The result of merging the EPTDs of
    \emph{la} and \emph{voit} by merging the nodes (1,2) and (1,5) with
    (2,2) and (2,4), respectively.}
\end{figure}

@@

We can see the EPTDs of both \emph{aucun} and \emph{ne} on
Figure~\ref{fig:ig-neg}. This example demonstrates two points of
interest to us.

@@

Finally, we explain the term $D((F_1 \limp ... \limp F_n \limp
G)^+)$. The $+$ operation recursively assigns positive and negative
polarities to all formula occurrences within $F_1 \limp ... \limp F_n
\limp G$. These polarities then drive the recursive definition of $D$,
which maps this polarized formula into a syntactic description. Since we
will not cover the theorem deeply in our treatise, we will settle for
knowing that the composition $D \circ +$ of the positive polarization
operation $+$ and the syntactic description producing operation $D$ maps
an IILL formula into a syntactic description such that the above-stated
theorem holds. Furthermore, this mapping can be easily constructed since
Perrier gives straightforward ways of computing the results of both
operations.

Perrier's theorem establishes a striking symmetry between the two
formalisms. Let us consider the case of parsing a sentence using a
lexicalized grammar in both formalisms. First, the sentence is tokenized
into wordforms and for each wordform, a lexical item is selected, be it
a typed constant in some abstract signature for ACGs or an EPTD for IGs.

In ACGs, we will try to take the typed constants and use them to produce
a term having some distinguished type $S$. Thanks to the Curry-Howard
correspondence, constructing a term having a given type is nothing more
than proving that term's type as a formula in the logic of our type
system. What this means is that our parsing problem boils down to
proving the IILL sequent $\tau_1, ..., \tau_n \vdash S$ (we omit the
inclusion of intuitionistic non-linear implications since they are not
relevant for parsing sentences). The sentence thus becomes parsable by
our grammar under the given lexical selection if and only if this
sequent is provable and furthermore, the proof of the sequent is our
desired syntactic structure in disguise.

@@

Secondly, a bigger problem is posed by the fact that $D \circ +$ is not
surjective. Perrier \cite{perrier2001intuitionistic} characterized the
set of syntactic descriptions that can be obtained as images of IILL
sequents using our transformation $D \circ +$, the set of \emph{IILL
  tree descriptions}, which is a proper subset of the set of all
syntactic descriptions. Specifically, in an IILL tree description, all
immediate dominance constraints go from negative nodes to positive
nodes, all (large) dominance constraints lead from positive nodes to
negative nodes and the root has a positive polarity.

@@

The PTDs of IGs include precedence constraints. We can see them as
serving two purposes, to both provide an ordering of the constituents
and to further constrain syntactic composition. In our translation to an
ACG, we could imagine handling the first role, that of providing word
order, by defining a lexicon from our abstract signature. Each constant
would combine its arguments in the order in which their corresponding
subtrees appeared in the ordered PTD. However, the second role of
precedence constraints, to establish further restrictions on syntactic
composition, would be lost.

@@

However, this translation is very lossy in terms of its discriminatory
power. Many of the constraints which were enforced by the original EPTD
are now ignored. We can imagine improving our strategy by using the
feature structures as our types instead of atomic categories, which
would not only ensure that e.g. the $sent\_type$ of the embedded clause
must be $decl$, but would also capture the information in the polarized
$funct$ feature so that the relative pronoun \emph{que} is not permitted
to extract subjects. There are still constraints that we would miss out
on though, such as the fact that the subject (2,3) of the clause (2,0)
must precede any embedded clauses (2,6)...(2,5), the fact that any such
embedded clauses must serve either the $modal$ or the $obj$ functions
and the fact that the embedded clause (2,5) must have a subject (2,4).
